Equalizer Programming Guide\\ {\vspace{-1ex}\footnotesize\sffamily\mdseries \htmladdnormallink{http://www.equalizergraphics.com/documents/Developer/ProgrammingGuide.pdf} {http://www.equalizergraphics.com/documents/ProgrammingGuide.pdf}}

### 2007-11-13 Developer/ProgrammingGuide/paper.tex
Update Quickstart Guide, fix some typos
[ ] May break build
[ ] Breaks existing applications
[ ] Bugfix
[ ] New Feature
[x] Cleanup
[ ] Optimization
[x] Documentation
- Implement OpenMP-based parallel image compression - Make object change manager configuration explicit - Programming Guide and documentation for the above
Existing eqNet::Object subclasses have now to use getChangeType()
instead of isStatic(), usePack() and/or isUnbuffered().


[ ] May break build
[x] Breaks existing applications
[ ] Bugfix
[x] New Feature
[x] Cleanup
[ ] Optimization
[ ] Documentation

\date{
  \vspace{2cm}
  \begin{figure}[ht]
    \hspace{-4.5cm}\includegraphics[width=22cm]{images/teaser.pdf}
  \end{figure}
  \vspace{1cm}
%  \textbf{RELEASE CANDIDATE}\\[\medskipamount]
  Version 1.0 for Equalizer 0.4\\[\medskipamount]
  \today
}

@@

\clearpage
\thispagestyle{empty}
\tableofcontents
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    \hline
  \end{tabularx}}
\thispagestyle{empty}
\clearpage

@@

\subsection{Execution Modeal and Thread Safety}

@@

@@

Versioned objects have to override \textsf{isStatic} to return
\textsf{false}, which indicates that they are versioned. The current
implementation has the following characteristics:
\begin{itemize}
\item Only the master instance of the object is writable, that is,
  \textsf{eqNet::Object::com\-mit} can be called only on the master
  instance to generate a new version.
\item Slave instance versions can only be advanced, that is,
  \textsf{eqNet::Object::sync( version )} with a version smaller than
  the current version will fail.
\item Newly mapped slave instance are mapped to the oldest available
  version.
\end{itemize}

@@

The \textsf{eqPly} frame data is layed out in one anonymous structure in
memory. It also does not track changes since it is relatively small in
size and changes frequently. Therefore, the instance and delta data are
the same and set in the constructor. The default implementation will
take care of the distribution of the data:

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \vspace{-8ex}\includegraphics[width=.382\textwidth]{images/DB.png}
  {\caption{\small\label{fDB}Destination view of an DB compound}}
\end{wrapfigure}
After the basic view setup, a directional light is configured, and the
model is positioned using the camera parameters from the frame data. The
camera parameters are transported using the the frame data to ensure
that all channels render a given frame using the same position.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/render.pdf}
  {\caption{\small\label{fRender}Main Rendering Loop}}
\end{wrapfigure}
The rendering loop maintains a list of candidates to render, which
initially contains the root node. Each candidate of this list is tested
for full visiblity against the frustum and range, and rendered if
visible. It is dropped if it is fully invisible or fully out of
range. If it is partially visible or partially in range, the children of
the node are added to the candidate list. \fig{fRender} shows a flow
chart of the rendering algorithm, which performs efficient view frustum
and range culling.

The actual rendering uses display lists or vertex buffer objects. These
OpenGL objects are allocated using the object manager. The rendering is
done by the leaf nodes, which are small enough to store the vertex
indices in a \textsf{short} value for optimal performance with VBO's.
The leaf nodes reuse the objects stored in the object manager, or create
and set up new objects if it was not yet set up. Since one object
manager is used per thread (pipe), this allows a thread-safe sharing of
the compiled display lists or VBO's across all windows of a pipe.

@@

Furthermore, the frame data transports the inherited range of the output
frame's compound. The range can be used to compute the assembly order of
multiple input frames, e.g., for sorted-blend compositing in volume
rendering applications.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/assembly.pdf}
  {\caption{\small\label{fAssembly}Hierarchy of assembly classes}}
\end{wrapfigure}
The offset of input and output frames characterizes the position of the
frame data with respect to the framebuffer, that is, the
\textbf{window's} lower-left corner. For output frames this is the
position of the channel with respect to the window.

>>>>

\date{
  \vspace{2cm}
  \begin{figure}[ht]
    \hspace{-4.5cm}\includegraphics[width=22cm]{images/teaser.pdf}
  \end{figure}
  \vspace{1cm}
%  \textbf{RELEASE CANDIDATE}\\[\medskipamount]
  Version 1.1.0 for Equalizer 0.4+svn\\[\medskipamount]
  \today
}

@@

\clearpage
\thispagestyle{empty}
\tableofcontents
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.0.1   & Nov 06, 2007 & Fixed Typo\\
    1.1     & Nov 12, 2007 & Adapted distributed objects\\
    \hline
  \end{tabularx}}
\thispagestyle{empty}
\clearpage

@@

\subsection{Execution Model and Thread Safety}

@@

\subsubsection{Change Handling}

The way changes are to be handled is determined by calling
\textsf{Object::getChangeType} during the registration of the master
version of a distributed object . The change type determines the memory
footprint and the contract for calling the serialization methods. The
following change types are possible:

\begin{description}
  \item[STATIC] The object is not versioned. The instance data is
    serialized whenever a new slave instance is mapped. No additional
    data is stored.
  \item[INSTANCE] The object is versioned, and the instance and delta
    data is identical, that is, only instance data is
    serialized. Previous instance data is saved to be able to map old
    versions.
  \item[DELTA] The object is versioned, and the delta data is typically
    smaller than the instance data. Both the delta and instance data are
    serialized and saved to map old versions.
  \item[DELTA\_UNBUFFERED] The object is versioned, and delta data is
    used to update slave instances. No data is stored, and no previous
    versions can be mapped. The instance data is serialized whenever a
    new slave instance is mapped.
\end{description}

@@

Versioned objects have to override \textsf{getChangeType} to indicate
how they want to have changes to be handled. The current
implementation has the following characteristics:
\begin{itemize}
\item Only the master instance of the object is writable, that is,
  \textsf{eqNet::Object::com\-mit} can be called only on the master
  instance to generate a new version.
\item Slave instance versions can only be advanced, that is,
  \textsf{eqNet::Object::sync( version )} with a version smaller than
  the current version will fail.
\item Newly mapped slave instance are mapped to the oldest available
  version.
\end{itemize}

@@

The \textsf{eqPly} frame data is layed out in one anonymous structure in
memory. It also does not track changes since it is relatively small in
size and changes frequently. Therefore, the instance and delta data are
the same and set in the constructor. Furthermore, the change type is
\textsf{INSTANCE}. The default \textsf{Object} implementation will take
care of the distribution of the data:

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/DB.png}
  {\caption{\small\label{fDB}Destination view of an DB compound}}
\end{wrapfigure}
After the basic view setup, a directional light is configured, and the
model is positioned using the camera parameters from the frame data. The
camera parameters are transported using the the frame data to ensure
that all channels render a given frame using the same position.

@@

The rendering loop maintains a list of candidates to render, which
initially contains the root node. Each candidate of this list is tested
for full visiblity against the frustum and range, and rendered if
visible. It is dropped if it is fully invisible or fully out of
range. If it is partially visible or partially in range, the children of
the node are added to the candidate list. \fig{fRender} shows a flow
chart of the rendering algorithm, which performs efficient view frustum
and range culling.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/render.pdf}
  {\caption{\small\label{fRender}Main Rendering Loop}}
\end{wrapfigure}
The actual rendering uses display lists or vertex buffer objects. These
OpenGL objects are allocated using the object manager. The rendering is
done by the leaf nodes, which are small enough to store the vertex
indices in a \textsf{short} value for optimal performance with VBO's.
The leaf nodes reuse the objects stored in the object manager, or create
and set up new objects if it was not yet set up. Since one object
manager is used per thread (pipe), this allows a thread-safe sharing of
the compiled display lists or VBO's across all windows of a pipe.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/assembly.pdf}
  {\caption{\small\label{fAssembly}Hierarchy of assembly classes}}
\end{wrapfigure}
Furthermore, the frame data transports the inherited range of the output
frame's compound. The range can be used to compute the assembly order of
multiple input frames, e.g., for sorted-blend compositing in volume
rendering applications.

@@

The offset of input and output frames characterizes the position of the
frame data with respect to the framebuffer, that is, the
\textbf{window's} lower-left corner. For output frames this is the
position of the channel with respect to the window.
